---
title: "OOI Alfresco data API example"
output: html_notebook
---

This notebook shows how to access CSV data from https://alfresco.oceanobservatories.org/ using R. 

This is a translation of the python method below:

```{python, eval = FALSE}
discrete_URL = ('https://alfresco.oceanobservatories.org/alfresco/webdav/OOI/'
                'Coastal%20Endurance%20Array/Cruise%20Data/Endurance-13_TN380_2020-07/'
                'Ship_Data/Water%20Sampling/Endurance-13_TN380_Discrete_Summary.csv')
r = requests.get(discrete_URL, auth=('guest', 'guest'))
times = ['Start Time [UTC]', 'CTD Bottle Closure Time [UTC]']
discrete = pd.read_csv(io.StringIO(r.content.decode('utf-8')), parse_dates=times)
discrete[96:]
```

We'll use {httr} for sending the request and recieving data, dplyr for munging, and ggplot2 for plotting.

```{r}
library(tidyverse)
library(httr)
```

Set the URL of interest.

```{r}
url <- 'https://alfresco.oceanobservatories.org/alfresco/webdav/OOI/Coastal%20Endurance%20Array/Cruise%20Data/Endurance-13_TN380_2020-07/Ship_Data/Water%20Sampling/Endurance-13_TN380_Discrete_Summary.csv'
```

Use `httr::GET` to send our request and return a response object. Note the very crafty username and password (guest, guest).

```{r}
req <- GET(url, authenticate("guest", "guest"))
```

Request objects are essentially augmented R lists.

```{r}
str(req)
```

We can look at the object by just printing it, get the headers with `headers()` and most useful, get the content or body of the request with `content()`.

```{r}
req
headers(req)
content(req)
```

Store the csv content in a data frame, select a cast and plot a profile! R is not a fan of verbose column names with spaces and punctuation, so we have to rename before plotting.

```{r}
df <- content(req)
head(df)
df %>% 
  filter(Cruise == "TN380",
         Station == "CE06",
         Cast == 1) %>% 
  rename(niskin = "Niskin/Bottle Position", chl = "Discrete Chlorophyll [ug/L]") %>% 
  ggplot(aes(niskin, chl)) +
  geom_point() +
  coord_flip()
```

